---
title: "Worshop, Part 8 - Filters & Joins"
author: "Axel Waldecker"
date: "6 August 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
library(kableExtra)
```

Reading in the `fish.csv` dataset.
```{r echo=T, message=FALSE}
fish <- read_csv(here("data","fish.csv"))
kelp_abur <- read_excel(here("data","kelp.xlsx"), sheet = "abur")
```

Filtering data of interest from the overall data.
```{r}
fish_mohk <- fish %>% 
  filter(site == "mohk")

head(fish_mohk)
```

Multi-condition querying: let's create a subset of all observations from **2016** *OR* where the species is **garibaldi**.
```{r}
fish_gar_2016 <- fish %>% 
  filter(year == 2016 | common_name == "garibaldi")

glimpse(fish_gar_2016)
```

Next, create a subset from the `fish` data frame, called `low_gb_wr` that only contains:

* *garibaldi* or *wrasse* observations **AND**
* *total_count* of 10 or less

```{r}
low_gb_wr <- fish %>% 
  filter(
    common_name %in% c("garibaldi", "rock wrasse") ,
    total_count <= 10
  )

head(low_gb_wr)
```

Create a new object called fish_it, starting from fish, that only contains observations if the common_name variable contains the string pattern “it”. What species remain?
```{r}
fish_it <- fish %>% 
  filter(str_detect(common_name, pattern = "it"))

unique(fish_it$common_name)

```

## Joining data together
Full join = safest otion, nothing is left out
```{r}
abur_fish_full <- kelp_abur %>% 
  full_join(fish, by = c("year", "site"))

View(abur_fish_full)
```

Left join = everything from original (=left) dataframe, plus any matches from the other one.
```{r}
abur_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year", "site"))

head(abur_fish_left)
```

Filtering, then joining in a sequence.
```{r}
my_fish_join <- fish %>% 
  filter(
    year == 2017, site == "abur" # only including observations for 2017 at 'abur'
  ) %>% 
  left_join(kelp_abur, by = c("year","site")) %>% 
  mutate(fish_per_frond = total_count / total_fronds)

head(my_fish_join)

kable(my_fish_join) %>% 
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = TRUE
  )
```
